---
title: "Assignment 5"
author: "Chadwick Reed"
date: "10/8/2020"
output:
   html_document:
    toc: TRUE
    toc_float: true
    toc_depth: 4
    message: FALSE
    results: 'hide'
    warning: FALSE
    tidy: TRUE
    error: FALSE
---


```{r}
library(tidyverse)
library(jtools)
library(interactions)
library(knitr)
```

### Description

In the last assignment, I used regression analysis to show the extent to which internet access (**ACCESS**) in the state of California is predicted by the following variables:


**BLD**: *Units per structure*
<br>
**NP**: *Number of persons per household*
<br>
**HPIP**: *Gross rent/ownership costs as a percentage of household income*


In this assignment, I add the following variables:


**BDSP**: *Number of bedrooms per household*
<br>
**HINCP**: *Household income*
<br>
**NPPB**: *Number of people per bedroom (created by dividing the **NP** variable by **BDSP***)

### Analysis

First, I will mutate BLD into fewer, shorter variable names in the interest of making my output as legible as possible. 
```{r}
hh_data_mod <- read.csv("hh_data_mod.csv") %>%
  mutate(struct_type = case_when(
    BLD_label=="2 Apartments" ~ "LT_10_apts",
    BLD_label=="10-19 Apartments" ~ "TenTo49apts",
    BLD_label=="One-family house detached" ~ "1SFhome",
    BLD_label=="3-4 Apartments" ~ "LT_10_apts",
    BLD_label=="Mobile home or trailer" ~ "MH_veh",
    BLD_label=="50 or more apartments" ~ "GT_49_apts",
    BLD_label=="20-49 Apartments" ~ "TenTo49apts",
    BLD_label=="5-9 Apartments" ~ "LT_10_apts",
    BLD_label=="One-family house attached" ~ "1SFhome",
    BLD_label=="Boat, RV, van, etc." ~ "MH_veh"))
```


Here, I convert all incompatible values to "NA" so that they'll be excluded from the sample during my analyses.
```{r}
hh_data_mod$BDSP[hh_data_mod$BDSP == -1] <-NA
hh_data_mod$HINCP[hh_data_mod$HINCP == -60000] <-NA
```


Next, I am simplifying the ACCESS variable so that there are only two results: **0** *(no internet)* and **1** *(has internet)*. My hope is that this facilitates the production of more meaningful results by my regression analysis.
```{r}
hh_data_mod$ACCESS[hh_data_mod$ACCESS == 2] <- 1
hh_data_mod$ACCESS[hh_data_mod$ACCESS == 3] <- 0
```


Finally, I create the variable 'NPPB' to measure number of persons per bedroom per household:
```{r}
hh_data_mod$NPPB <- hh_data_mod$NP/hh_data_mod$BDSP
```


With that done, here are my models:
```{r}
model1 <- lm(ACCESS ~ struct_type  + NP  + HPIP ,
            data = hh_data_mod)

summary(model1)
```


```{r}
model2 <- lm(ACCESS ~ struct_type  + NP  + HINCP ,
            data = hh_data_mod)

summary(model2)
```


```{r}
model3 <- lm(ACCESS ~ struct_type  + BDSP  + HINCP ,
            data = hh_data_mod)

summary(model3)
```















