---
title: "Assignment 1"
author: "Chadwick Reed"
date: "9/9/2020"
output: html_document
---
### Analyzing the Relationship Between Internet Access and Interstructural Population Density

<br>

This dataset is assembled with the intention of analyzing the extent to which internet access is a function of interstructural population density. It is well documented that internet access increases with regional population density: A downtown city center will undoubtedly have greater access than a rural farm town. This dataset attempts to test (or, rather, provide a means to test) whether the relationship between internet access and density holds true at the micro-scale; by assessing individual households across the state of California.

<br>

The following variables are included in the dataset:
<br>
BLD: Units per structure
<br>
NP: Number of persons per household
<br>
HINCP/OCPIP: Gross rent/owner costs as a percentage of household income
<br>
ACR: Lot size
<br>
TEN: Tenure
<br>
ACCESS: Internet access

<br>
<br>

This dataset includes a total of 378,817 observations.

<br>

```{r load libraries,message=FALSE, results='hide'}
library(tidycensus)
library(tidyverse)
```
```{r}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```
```{r results='hide'}
hh_data <- get_pums(variables = c("BLD", 
                                  "NP", 
                                  "HINCP",
                                  "GRPIP",
                                  "OCPIP",
                                  "ACR",
                                  "TEN", 
                                  "ACCESS",
                                  "HISPEED"),
                    state = "CA",
                    year = 2018,
                    survey = "acs1")
```
```{r}
hh_data
```
